The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
Proof-read section 2.1, 2.2

Literature on anthropomorphism is quite diverse. We would like to point out that there exist different and partly conflicting definitions of the term anthropomorphism, and that in general, the \textit{terminology} is not clear. Probably as a result of this, there also exist different understandings of the \textit{phenomenon} that the term `anthropomorphism' describes.
 In this section we will first give an overview on the diversity of terms and then try to establish a common ground in the understanding of anthropomorphism (the phenomenon).

\subsection{Overview of the diversity of terms}

Originally, the term \textbf{anthropomorphism} comes from the Greek \textit{anthropos} for `man' (or `human') and \textit{morphe} for `form / structure' (or `shape'). Anthropomorphism is generally understood as people's tendency to perceive a non-human agent as if it had some human characteristics or traits. How a person perceives an agent is crucial, as it further determines how she relates to, and behaves towards the agent, or how she may behave in light of this agent \citep{epley_when_2008}.
As such, anthropomorphism can be seen as a special kind of relationship between a human interacting with a non-human agent. An anthropomorphic relationship is often characterized by the fact that the human attributes human-unique characteristics to the non-human agent. Since anthropomorphism becomes observable through these attributions, some definitions claim that anthropomorphism is not the \textit{perception} but rather the \textit{attribution} of human qualities to non-human agents. Ascribing thoughts, emotions, and intentions are given as examples, and as the most prominently anthropomorphized non-human agents animals and robots are mentioned \citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism there also exists the term \textbf{animism}, which can be defined as the attribution of life to the nonliving. The study of animism was pioneered by Piaget in the 1920s with his work on \textit{The Child's Conception of the  World}. He found that children tend to regard many inanimate objects as capable of sensations, emotions and intentions, and called this view `animism'. Since then, children`s inferences about the world have been studied intensively, and there is an agreement that children`s views and concepts of living, animal, and plant evolve as they grow up. Often synonymously used to animism is the term \textbf{zoomorphism}, which similarly applies to the cases when non-lifelike instances / artifacts are associated with animalistic attributes, excluding human-specific traits.
As `being alive' is not a human-unique property but applies to all living beings, the attribution of life may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions of animacy, is to look at paraphrases of anthropomorphism. That is to say, 
anthropomorphism can alternatively be understood as \textbf{humanization} or 
\textbf{personification} of anything other than a human being.\footnote{There exist also the phenomenons inverse to anthropomorphism, which can be found in the literature as \textbf{dehumanization} \citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism} \citep{caporael_anthropomorphism_1986}. However, we do not treat this here.} 
Robots can both be animized and anthropomorphized, but often the difference is not made.

@@

Naturally, conflicts in how terms are used lead to a diversity of understandings of the phenomenon which is described.
One reason for this might be that anthropomorphism
is a phenomenon that is studied in very different domains that might each
integrate their own understanding, or simply use the same term to refer to different things.
The conflict when reading about anthropomorphism is, that despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. It is even
argued that these terms might be misused \citep{duffy_anthropomorphism_2002,epley_when_2008}.
For instance, some researchers refer to \textit{``the
robot's level of anthropomorphism''} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they understand that a system or artifact itself does not \textit{``contain anthropomorphism''} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such \hl{(to which we will come back in the next section)}.

Another paradox is that, while human's tendency to anthropomorphize is commonly observed, the phenomenon is 
at the same time still rather poorly understood
\cite{epley_seeing_2007}. The phenomenon itself has been found to be very complex though sometimes
subtle and as such hard to study \cite{duffy_anthropomorphism_2002,epley_when_2008}. Further, it is not clear how to operationalize and measure anthropomorphism. We would like to discuss the challenges of studying and measuring anthropomorphism toward the end of this article, in Section \ref{sec:measuring}.

%Anthropomorphism is used in different senses throughout different disciplines 
%\citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings. 
Depending on which definition of anthropomorphism is applied, not everything 
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.
One practical example is that, it is not always clear whether a person (may it be an adult or 
a child) who gives and anthropomorphic response, believes that the other agent really 
thinks or acts like a human or whether she is using anthropomorphism in a metaphoric 
sense \citep{leeds_childrens_1992}.
Thus, we need to be clear about what we mean when talking about anthropomorphism.

Several attempts have been made to clarify what anthropomorphism is (the phenomenon).
From a psychology point of view \cite{epley_when_2008} try to clarify things by explaining what anthropomorphism is \textit{not}. For instance, in their understanding, anthropomorphism does not include behavioral descriptions of observable actions but it requires going beyond what is directly observable. It is further proposed that anthropomorphism does not merely entail \textit{animism}, as animate life is not a \textit{uniquely} human-like characteristic.\footnote{For the whole list and more details, the reader may refer to the original work \citep{epley_when_2008}.}

@@

As human-likeness itself is a huge dimension, some researchers argue that there might not just be one single type of anthropomorphism but different levels or gradations. For instance, \cite{persson_anthropomorphism_2000} proposes to understand anthropomorphism as a multi-layered phenomenon, and suggests that it exists of different levels, namely primitive categorization, primitive psychology, folk-psychology, traits, social roles, and emotional anthropomorphism. Correspondingly, \cite{ruijten_introducing_2014} propose to understand anthropomorphism as a continuum ranging from low to high human likeness, and to measure it as such (see Section \ref{sec:measuring}).

>>>>

Literature on anthropomorphism is quite diverse. We would like to point out that there exist different and partly conflicting definitions of the term anthropomorphism, and that in general, the \textit{terminology} is not clear. Probably as a result of this, there also exist different understandings of the \textit{phenomenon} that the term `anthropomorphism' describes.
 In this section we will first give an overview on the diversity of terms and
 then try to establish a common ground in the understanding of anthropomorphism
 as a psychological phenomenon.

 \subsection{The lexical field of anthropomorphism}

Originally, the term \textbf{anthropomorphism} comes from the Greek
\textit{anthropos} for ``man'' (or ``human'') and \textit{morphe} for `form /
structure' (or `shape'). Anthropomorphism is generally understood as people's
tendency to perceive a non-human agent as if it had some human characteristics
or traits. How a person perceives an agent is crucial, as it further determines
how he/she relates to, and behaves towards the agent, or how he/she may behave in
light of this agent \citep{epley_when_2008}. An anthropomorphic relationship is
often characterized by the fact that the human attributes human-unique
characteristics to the non-human agent. Since anthropomorphism becomes
observable through these attributions, some definitions claim that
anthropomorphism is not the \textit{perception} but rather the
\textit{attribution} of human qualities to non-human agents.  Ascribing
thoughts, emotions, and intentions are given as examples, and as the most
prominently anthropomorphized non-human agents, animals and robots are mentioned
\citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism, there also exists the term \textbf{animism}, which
can be defined as the attribution of life to the nonliving. The study of animism
was pioneered by Piaget in the 1920s with his work on \textit{The Child's
Conception of the  World}. He found that children tend to regard many inanimate
objects as capable of sensations, emotions and intentions, and called this view
``animism''. Since then, children's inferences about the world have been studied
intensively, and there is an agreement that children's views and concepts of
living, animal, and plant evolve as they grow up. Often synonymously used to
animism is the term \textbf{zoomorphism}, which similarly applies to the cases
when non-lifelike instances / artifacts are associated with animalistic
attributes, excluding human-specific traits.  As ``being alive'' is not a
human-unique property but applies to all living beings, the attribution of life
may not be labeled as anthropomorphism (in a narrow understanding of it).
Another possibility to illustrate how anthropomorphism differs from attributions
of animacy is to look at paraphrases of anthropomorphism. That is to say,
anthropomorphism can alternatively be understood as \textbf{humanization} or
\textbf{personification} of anything other than a human being\footnote{There
exist also the phenomenons inverse to anthropomorphism, which can be found
in the literature as \textbf{dehumanization}
\citep{haslam_dehumanization:_2006}, or \textbf{mechanomorphism}
\citep{caporael_anthropomorphism_1986}.}. Robots can
both be animized and anthropomorphized, but often the difference is not made.

@@

Naturally, conflicts in how terms are used lead to a diversity of understandings
of the phenomenon which is described.  One reason for this might be that
anthropomorphism is a phenomenon that is studied in very different domains that
might each integrate their own understanding, use the same term to refer to
different things.  Despite the fact that there is no commonly accepted
definition, the terms \textit{anthropomorphism}, \textit{anthropomorphic} or
\textit{human-like} are often used as if their meanings were clear and agreed
upon \citep{persson_anthropomorphism_2000}. It is even argued that these terms
might be misused \citep{duffy_anthropomorphism_2002,epley_when_2008}.  For
instance, some researchers refer to \textit{``the robot's level of
anthropomorphism''} \citep{bartneck_is_2007}, whereas others disagree to such a
statement because they understand that a system or artifact itself does not
\textit{``contain anthropomorphism''} but only gives rise to the process of
anthropomorphizing in a given user and situation. Consequently,
\cite{persson_anthropomorphism_2000} conclude that anthropomorphism emerges in
the \textit{interaction} between the technology and the user, and has to be
understood as such. We come back on this point in the next section.

Another paradox is that, while human's tendency to anthropomorphize is commonly
observed, the phenomenon is at the same time still rather poorly understood
\citep{epley_seeing_2007}. The phenomenon itself has been found to be very
complex, sometimes subtle and as such hard to study
\citep{duffy_anthropomorphism_2002,epley_when_2008}. Besides, it is not clear
how to operationalize and measure anthropomorphism (we discuss this question
toward the end of this article, in Section \ref{sec:measuring}).

%Anthropomorphism is used in different senses throughout different disciplines 
%\citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings. 
Depending on which definition of anthropomorphism is applied, not everything 
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.
For example, it is not always clear whether a person (may it be an adult or 
a child) who gives an anthropomorphic response, believes that the other agent really 
thinks or acts like a human or whether he/she is using anthropomorphism in a metaphoric 
sense \citep{leeds_childrens_1992}.
This call for explicitely stating what we mean when talking about anthropomorphism.

Several attempts have been made to clarify what the psychological phenomenon
\emph{anthropomorphism} actually is.  From a psychology point of view,
\citet{epley_when_2008} explain what anthropomorphism is \textit{not}. For
instance, in their understanding, anthropomorphism does not include behavioral
descriptions of observable actions but it requires going beyond what is directly
observable. They also suggest that anthropomorphism does not merely entail
\textit{animism}, as animate life is not a \textit{uniquely} human-like
characteristic\footnote{For the whole list and more details, the reader may
refer to the original work \citep{epley_when_2008}.}.

@@

As human-likeness itself is a huge dimension\fxwarning{SL: huge dimension?},
some researchers argue that there might not just be one single type of
anthropomorphism but different levels or gradations. For instance,
\citet{persson_anthropomorphism_2000} proposes to understand anthropomorphism as
a multi-layered phenomenon, namely \emph{primitive categorization},
\emph{primitive psychology}, \emph{folk-psychology}, \emph{traits}, \emph{social
roles}, and \emph{emotional anthropomorphism}. Correspondingly,
\citet{ruijten_introducing_2014} propose to understand anthropomorphism as a
continuum ranging from low to high human likeness, and to measure it as such.
